Bootstrap learning via modular concept discovery

### 2014-06-08 old/drafts/ijcai2013/figures/4sqtree.tex
Moved everything into an a new 'old' directory.

\documentclass{article}

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{amsfonts}
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{float}
\usepackage{circuitikz}
\usepackage{graphicx}
\usepackage[font=small]{subcaption}
\usepackage[font=small]{caption}
\usepackage{alltt}
\usepackage{dashrule}
\usepackage{array}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}

\usepackage{times}
\pdfinfo{
  /Title (Bootstrap learning via modular concept discovery)
  /Author (Eyal Dechter, Jon Malmaud, Ryan P. Adams, Joshua B. Tenenbaum)
}

\newtheorem{definition}{Definition}
\DeclareMathOperator*{\argmax}{arg\,max} 
\DeclareMathOperator*{\argmin}{arg\,min}

\title{Bootstrap learning via modular concept discovery}

\author{Eyal Dechter \\
MIT\\
USA \\
edechter@mit.edu
\And
Jon Malmaud \\
MIT \\ 
USA \\
malmaud@mit.edu
\And 
Ryan~P.  Adams \\
Harvard University\\
USA \\
rpa@seas.harvard.edu
\And
Joshua~B. Tenenbaum \\
MIT\\
USA \\
jbt@mit.edu
}

\begin{document}

\maketitle
\begin{abstract}
%\small
 Suppose a learner is faced with a domain of problems about which it
 knows nearly nothing. It does not know the distribution of problems,
 the space of solutions is not smooth, and the reward signal is
 uninformative, providing perhaps a few bits of information but not
 enough to steer the learner effectively. How can such a learner ever
 get off the ground? A common intuition is that if the solutions to
 these problems share a common structure, and the learner can solve
 some simple problems by brute force, it should be able to extract
 useful components from these solutions and, by composing them,
 explore the solution space more efficiently. Here, we formalize this
 intuition, where the solution space is the space of typed functional
 programs and the gained information is stored as a stochastic grammar
 over programs. We propose an iterative procedure for exploring such
 spaces: in the first step of each iteration, the learner explores a
 finite subset of the domain, guided by a stochastic grammar; in the
 second step, the learner compresses the successful solutions from the
 first step to estimate a new stochastic grammar. We test this
 procedure on symbolic regression and Boolean circuit learning and
 show that the learner discovers modular concepts for these
 domains. Whereas the learner is able to solve almost none of the
 posed problems in the procedure's first iteration, it rapidly becomes
 able to solve a large number by gaining abstract knowledge of the
 structure of the solution space.

\end{abstract}

\section{Introduction}
A central question for AI is how to create intelligent agents without
building in vast amounts of pre-specified knowledge that is tailored
to one domain or another. Any attempt to solve this problem must
contend with the fact that the less an agent knows about solving
problems in a domain, the worse its attempts are, and, therefore, the
less informative they are about how to solve subsequent problems.

For example, imagine you know nothing about electronics but are
supplied with a collection of circuit elements and asked to build a
useful device: your initial attempts will be uniformly
ineffective. If, on the other hand, you know the basics, you are far
more likely to generate devices that are somewhat successful and on
whose basis you might improve your skills.

Like many of the problems in AI, this one can be cast as a search
problem: the goal is to find a circuit wiring procedure that maximizes
certain criteria.  There are many search algorithms, but they typically
exploit local or topological structure in the search space. For
example, local search algorithms move from one candidate solution to a
neighboring one. Heuristic search techniques take advantage of
topological structure to prune off large parts of the solution space.

Domain specific knowledge provides this structure. In local search, it
defines the similarity metric between solutions; in heuristic search
it provides the heuristic. Here we ask how to bootstrap domain
specific knowledge and extract structure from the space of possible
solutions by attempting to solve multiple related problems in a given
domain at once. Our approach is motivated by the observation that many
interesting problems -- such as discovering scientific theories,
developing new technologies, learning to play a new instrument -- have
solutions that are natrually specified as programs.

Our proposed solution is thus inspired by the programmer who -- having
reused subroutines to solve similar simple tasks -- creates a function
library to help manage the complexity of larger projects. The
E.C. algorithm, which we introduce here, builds a library of reusable
program components and places a distribution over these, effectively
learning from simple tasks a search space structure that enables it to
solve more complex tasks.

In this work, we try to see how far this idea of reusable function
definition can get us. We ask the following question: can an AI system
discover the structure latent in the solutions to multiple related
problems and thereby bootstrap effective exploration of the search
space? Can discovering modular program components transform a problem
from one in which search is intractable into a problem in which it
becomes increasingly feasible with experience?

Concretely, the E.C. algorithm works as follows. It begins with a small
library of program primitives and a distribution over programs
synthesized from this library. In each iteration, it generates
programs in order of highest probability according to the library
learned in the previous iteration. From the discovered solutions, it
finds new reusable functions that compress these solutions, and it
re-estimates the distribution over programs by weighting these new
functions according to their frequencies. In using compression as a
guide to the choice of representation, we instantiate the idea that
good representations are those which minimize the description length
of typical elements in the domain.

We show that this iterative approach allows us to achieve very high
performance in tasks where just searching according to the initial
distribution over program primitives clearly fails.

\section{Related Work}
The idea of discovering and using reusable subcomponents is part of an
old tradition in AI. For instance, it is one of the central themes in
Herb Simon's 1969 ``The Sciences of the
Artificial''~\shortcite{simon1996sciences}. Rendell's 1985
paper~\shortcite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
Induction Using Layered Information Compression: Tractable Feature
Formation in Search,'' presented within a classical search paradigm
the idea that compression provides a heuristic for constructing good
representations. Within the Inductive Logic Programming (ILP)
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
was also applied in the multitask
setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
learning functional programs and believe that the modularity and
compositionality afforded by this representation is necessary for
these ideas to be successful.

Genetic programming
(GP)~\cite{DBLP:conf/foga/Koza92,DBLP:conf/ices/KozaBAK96} has
tackled the problem of program learning and has explored the notion
that reusable functions are helpful. This work relies, however, on
stochastic local search over the space of programs, and automatically
discovering useful subroutines is seen as a helpful heuristic. More
recently, Liang et al.~\shortcite{liang10programs} used a stochastic
grammar over combinatory logic to induce shared structure in
multi-task program induction problems. We find this latter
representation compelling and use it here, but we see both this and
genetic programming as tackling the question of how to benefit from
shared structure in situations where local search might be successful
on its own. Such problems rely on domain knowledge -- whether in the
form of a fitness function or a likelihood function -- to provide
local structure to the search space.

But, as mentioned before, the problems we want AI systems to solve
often lack this kind of built in structure. Here we attempt to learn
functional programs to solve those kinds of problems.  \\

%% AI and machine learning algorithms commonly take advantage of the
%% relationship between a performance metric (error, reward, fitness,
%% etc.) and the local or topological structure of the solution
%% space. Whether in discrete or continuous domains, the key is to design
%% such a relationship and write a algorithm that either moves in good
%% directions or prunes off large parts of the solution space. Whether or
%% not such a relationship exists depends on the representation, and much
%% work has gone into finding good representations for problem
%% solving. Recent work, especially in machine learning, has focused on
%% ``feature learning,'' with the explicit aim of discovered good
%% representational building blocks in vector-valued solutions spaces.

%% Many of the most natural solution representations for intelligent
%% agents, however, are not naturally expressed as vector-valued
%% spaces. Consider the problems of discovering scientific theories,
%% inventing new technologies, communicating with others, or executing a
%% complex motor routine (such as dancing or playing an instrument). The
%% solutions to these problems are most naturally expressed as programs,
%% that is, as the composition of modular components whose semantics are
%% determined by the actions of one component on another.

%% How can we learn good solution representations for these kinds of
%% tasks without building in the kind of domain knowledge that gives rise
%% to a useful relationship between representation and performance? How
%% can we get traction on a set of problems when small tweaks to
%% solutions give rise to wildly varying performances in an unpredictable
%% manner, or when, for most problems, most of the solutions fail in a
%% uniformly bad way?

%% A popular idea is that good representations are those that encode the
%% data space in an efficient way, that is, that enable compression of
%% the data~\cite{}. For program representations, this most naturally
%% corresponds to defining reusable components so that the total number
%% of unique expressions for a set of solutions is minimized (i.e., this
%% is directly akin to the subroutine libraries programmers use to manage
%% the complexity of their code).

%% In this work, we try to see how far reusable subroutine definition can
%% get us on its own. Our question is, Can discovering modular components
%% transform a problem from one in which brute-force search will not work
%% to a problem in which it does?

%% To emphasize an impoverished relationship between reward signals and
%% program representations, we make our tasks all-or-none: a solution
%% either solves the problem or it does not, and partial solutions do not
%% get any credit.

%% The intuitive approach to applying this idea is simple: if solutions
%% to problems in the domain share common structure, and there exist
%% problems whose solutions are simple enough to find by brute force,
%% then we should be able to extract a distribution over solution
%% subcomponents from these simple cases that allows us to run subsequent
%% searches more efficiently.

%% More concretely, our algorithm works as follows: each iteration, it
%% first generates programs in order of highest probability according to
%% a subroutine distribution learned in the previous iteration. From the
%% discovered solutions, it finds new reusable subroutines that compress
%% these solutions and estimates their probabilities.

%% We show that this iterative approach allows us to achieve very high
%% performance in tasks where just searching according to the initial
%% distribution over program primitives clearly fails.

%%  \hline

%% AI and machine learning algorithms commonly take advantage of the idea
%% that local structure in the solution space is related to local
%% structure in the outcome. In continuous domains, learning algorithms
%% algorithms use error gradients to guide search. In discrete domains,
%% whether in heuristic or local search paradigms, knowledge about the
%% solution space topology and its relation to reward and cost metrics
%% enables pruning, reasoning, and moving in the solution space more
%% efficiently. However, choosing the representations that have
%% beneficial structural properties requires a lot of general and domain
%% knowledge, and this what researchers do and what they cannot leave up
%% to their algorithms.

%% What is a good strategy for an agent when its representation of the
%% solution space is not smooth with respect to the outcome? Many hard
%% learning problems have this property, especially those in which the
%% solution space is most naturally represented as a set of
%% programs. Consider the problems of discovering scientific theories,
%% inventing new technologies, learning to cook, or learning to
%% communicate. These are problems in which 1) solutions are best
%% described as programs (i.e. compositions of modular subroutines at
%% various levels of abstraction), 2) most of the solution space is bad
%% in a relatively uninformative way (i.e. it just doesn't work), and
%% thus 3) outcomes are not informative about how you are doing unless
%% you are already quite close to a solution.

%% Intuitively, the way to solve such problems is to find solutions to
%% simple cases in the domain and extract conceptual building blocks that
%% can be composed to scale up to complex cases. With experience, the
%% agent may be able to discover local structure in the solution space,
%% and utilize more traditional AI and machine learning techniques. In
%% this work, we will be focusing on the first phase of this learning
%% problem.

%% %% Of course, every learner

%% %% Choosing the wrong representation, in
%% %% which local structure is not apparent, is often the difference between
%% %% a successful learner and an unsuccessful one. In many respects, this
%% %% may be the kind of situation that adult humans face when confronting
%% %% new domains: we have a lot of general knowledge of successful
%% %% representations, useful strategies, etc.

%% %% But suppose that you have very little knowledge of a domain. Your
%% %% intial representation of the domain is poor: ``nearby'' points in the
%% %% solutions space ellicit wildly different outcomes in unpredictable
%% %% ways. Or perhaps nearby points ellicit the same outcome and its not a
%% %% good one, so you don't know which way to go. Many learning problems,
%% %% especially those whose solutions are best described as programs, have
%% %% these qualities: imagine learning to synthesize new sentences

%% %% teness. We choose the representations our learners use
%% %% and these representations generally Imagine you are trying to learn to
%% %% bake via trial and error. You know what kind of ingredients are used
%% %% in baking (flour, water, butter, sugar, yeast, baking soda, etc.) and
%% %% you know a bit about how to combine these ingredient (mix, whisk,
%% %% etc.) and how to apply heat (oven). But baking is a notoriously
%% %% fragile process: ratios of ingredients, the order of mixing, cooking
%% %% temperatures, resting times, all make a huge difference, and the
%% %% outcome of the baking process is mostly all or none. Either the
%% %% outcome is what you expected or quite unlike it. Even if you do create
%% %% something similar to your intent, it is difficult to assign blame for
%% %% the outcome to any one step or decision along the way.

%% %% Many difficult learning problems have this structure -- learning
%% %% language, developing technologies and scientific theories, learning to
%% %% play instruments or execute other complex motor patterns. These are
%% %% problems in which 1) there is a large set of similar tasks; 2) their
%% %% solutions are best described as programs; 3) the solution space is
%% %% not smooth with respect to the outcome; 4) and the outcome is generally
%% %% uninformative unless you are already pretty close to a solution.

%% %% Of course, knowledge is not acquired wholly by trial and error. But in
%% %% many domains, there is at least initially such a period of ignorance,
%% %% and the question we are trying to tackle here is how you ever get off
%% %% the ground.

%% I need help here.

%% %% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
%% %% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
%% %% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
%% %% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
%% %% possible, as efficiently as possible. That is, suppose that there is a
%% %% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
%% %% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
%% %% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
%% %% before the first expression in $O$ to solve $t_2$.

%% %% We motivate a solution to this problem by considering a heirarchical
%% %% Bayesian formulation of the task distribution. Suppose we believe that
%% %% there is some underlying structure to $p_t(\cdot)$ that is explained by
%% %% the fact that for each task $t$ there is a latent expressions $e \in
%% %% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that

%% %% \begin{align}
%% %% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
%% %% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
%% %% \end{align}

%% %% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
%% %% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
%% %% expressions in this sum, there is some expression $e^*$ such that
%% %% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
%% %% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]

%% %% This suggests a way to find the ordering $O$: infer a stochastic
%% %% grammar $G$, under the assumption that $G$ favors one solution to a
%% %% given task over any other, and order the elements of $\mathcal{L}$
%% %% according to $G$.

%% %% Thus, our basic procedure will be to search for a single ``good''
%% %% solution for as many tasks as we can, and to use those solutions to
%% %% estimate a stochastic grammar. Our search will be performed in the
%% %% order of the most recently inferred grammar. So as our ordering
%% %% becomes more efficient and we can solve more tasks, we learn better
%% %% stochastic grammars.

%% %% Formally then, our goal is to solve the following minimal feedback
%% %% multi-task program induction problem: suppose we are given set of
%% %% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
%% %% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
%% %% programs in our language.  We say that $ p \in \mathcal{L}$
%% %% \emph{solves} $t_i$ if $ t_i(p) = 1$. Our goal is to find a set of
%% %% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that

%% %% \[
%% %% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
%% %% \]

%% Based on the motivation set out above, we present an iterative
%% strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
%% distribution over the primitives from $\mathcal{L}$ and a integer
%% \emph{frontier size} $N$, we perform two steps in each iteration.  In
%% the first step, we explore the $N$ highest probability programs
%% according to $\mathcal{G}$. In the second step, we compress the
%% programs that solve any of our tasks, and we use that compression to
%% define a new grammar. Repeating these two steps, we effectively
%% reparameterize our search space to include programs that are more likely
%% to be solutions to our tasks.

The remainder of this paper will include two major sections. In the
first, we describe in detail the E.C. algorithm and design choices we
make in implementing it. We will then present results from two
experiments in the domains of symbolic regression and Boolean
function learning.

%% \section{}
%% Let $\mathcal{F}$ be a family of distributions over $\mathcal{L}$ such
%% that for any $\mathcal{D} \in \mathcal{F}$ we can a) evaluate the
%% probability of program $p$, $P_D(p)$, and we can b) enumerate elements
%% from $D$ in descending order.

%%  To motivate a solution to our problem, we begin with the hierarchical
%% Bayesian formulation of multi-task program induction found in Liang et
%% al.~\cite{liang10programs}. The general idea there is to estimate
%% solution programs for a set of related problems by inferring the posterior
%% distribution over solutions under the assumption that all the solution
%% programs are generated from a common latent stochastic grammar. By
%% jointly inferring solutions and the latent grammar, we find a set of
%% solutions that favors reuse of subcomponents.

%% While this framework is appealing in its elegance and Bayesian
%% interpretation, it cannot be straightforwardly adapted to the
%% situation we are considering here because their proposed MCMC
%% algorithm relies on a local log-likelihood signal, the very kind that
%% we are assuming we do not have access to.

%% To begin, we provide a high level overview of the algorithm and the
%% problem that it solves. The problem setup requires a representation
%% language for programs and a family of distributions over expressions
%% in that language, but we leave these details for the following section

\section{Extracting concepts from programs}

Our goal is to solve the following minimal feedback multi-task program
induction problem: suppose we are given a set of \textbf{tasks}
$T=\{t_k\}_{k=1}^K$ where each task is a function $t_k : \mathcal{L}
\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of
expressions in our language.  We say that expression $ e \in
\mathcal{L}$ \textbf{solves} $t_k$ if $ t_k(e) = 1$. Our goal is to
solve as many of the tasks in $T$ as we can.

We propose the \textbf{E.C. algorithm}, an iterative procedure to solve this
problem. It will be useful in presenting the algorithm to define the
following terminology: a \textbf{frontier} is the finite set of
expressions that the algorithm considers in any one iteration. The
\textbf{frontier size} $N$ is the prespecified number of elements in
the frontier. We will say that a task is \textbf{hit} by the frontier
if there is a expression in the frontier that solves it.

The E.C. algorithm maintains a distribution $\mathcal{D}$ over
expressions in $\mathcal{L}$. At each iteration, it
\begin{enumerate}
\item \textbf{explores} the frontier -- enumerating the $N$ most probable expressions
  from the current distribution $\mathcal{D}$ -- and
\item \textbf{compresses} the hit tasks in the frontier to estimate a new
  distribution.
\end{enumerate}

\subsection{Representing typed functional programs as binary trees}

Following~\cite{liang10programs} and~\cite{briggs2006functional}, we use a
simply typed combinatory logic -- a variable-free subset of the
polymorphic simply-typed lambda calculus~\cite{Pierce_2002} -- as our
program representation language. In short, the combinatory logic
introduces a \emph{basis} of several primitive \emph{combinators} such
that any function can be written as applications of the basis
combinators. Some common basis combinators are defined as follows:
\begin{align}
\mathbf{I}\; x &\rightarrow  x & \text{ (identity) }\\
\mathbf{S}\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x)\\ 
\mathbf{C}\; f\; g\; x\; &\rightarrow (f\; x)\; g  \\ 
\mathbf{B}\; f\; g\; x\; &\rightarrow f\; (g\; x) & \text{ (composition) }
\end{align}

The basis combinators are theoretically enough to express any Turing
computable function, but we will assume that our language has a
variety of primitives and, together with the basis combinators, we
call these the \textbf{primitive combinators}. Note that by default we consider
all primitives to be in their ``curried'' form (e.g. a function like
``+'' adds two numbers $x$ and $y$ by being first applied to $x$,
returning the function $(+\;x)$ and then being applied to $y$,
returning $(+\,x\,y)$\;).

The basis combinators can themselves be expressed in the lambda
calculus (with definitions following directly from the equations
above). The lambda calculus has two basic operations -- application
and abstraction -- but in using the
combinatory logic we sequester uses of the abstraction operation
inside the combinators. In doing so, we have replaced the variable
binding of the $\lambda$ operator with the variable \emph{routing} of
these basis combinators. Our representation thus becomes
variable-free. See \cite{liang10programs} for a more
detailed discussion of this routing interpretation.

Using combinatory logic is very convenient for program
synthesis~\cite{briggs2006functional}: since every expression is the
application of one expression to another -- with this recursion
bottoming out at the primitive combinators -- each program is a binary
tree. Most importantly, any subtree is itself a well-formed
expression; this is not the case in the lambda calculus, since
abstraction introduces long range dependencies between the $\lambda$
operator and the variables to which that operator refers. In the
lambda calculus, then, a subtree might have free variables, not bound
to any enclosing $\lambda$.

As a simple example of our representation, consider the squaring
function $\lambda x. * x\; x $. Using two of the basis combinators
above, we can write the squaring function in combinatory logic as $S\; *\;
I$ (taking $*$ as a primitive), where application associates to the
left, so we drop the parentheses. When we apply this combinator to a
value $x$, the action of the combinator is defined as $S \;*\; I\;
x~\rightarrow~(*\; x)\; (I\; x)~\rightarrow *\; x \; x$.

We can extend this representation with a simple polymorphic type
system~\cite{Pierce_2002}. In this representation, a type is either a
type primitive (e.g. reals, integers, Booleans, etc.), a type variable
(e.g. $\sigma$), or a function type $ \tau_1 \rightarrow \tau_2$ of
functions from source type $\tau_1$ to target type $\tau_2$. Any combinator
can be represented as a binary tree~(Figure~\ref{fig:clbintree}) whose
leaves are typed primitive combinators and whose interior nodes
represent typed applications of combinators (we annotate interior
nodes with their types).

\begin{figure}
%%  \centering
  \begin{subfigure}[Before]{0.45\linewidth}
    \begin{tikzpicture}[scale=.8,
        transform shape, level/.style={sibling distance=5mm/#1, font=\scriptsize}
        ]
      \node [font=\scriptsize] (top) {$\mathbb{R}$}
      child {node [above=.3cm, left=.1cm, text width=1.8cm] (a) {$(\mathbb{R} \rightarrow \mathbb{R}) \rightarrow \mathbb{R} 
          \rightarrow \mathbb{R}$}
        child {node [left=.5cm, below=1cm, text width=3.0cm] 
          {$ \mathbf{S} : (t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
            \rightarrow$ \\ $t_2 \rightarrow t_0$}}
        child {node [right=0cm, text width=2cm] {$ \mathbf{*}: \mathbb{R} \rightarrow \mathbb{R} \rightarrow \mathbb{R}$}}
      }
      child {node [below right=-.5cm and .2cm] (b) {$\mathbb{R} \rightarrow \mathbb{R}$}
        child {node [above=.5cm] {$ \mathbf{I}: t_0 \rightarrow t_0 \rightarrow t_0$}}
      };
    \end{tikzpicture}
    \caption{    \label{fig:clbintree}}
  \end{subfigure}
  \begin{subfigure}[After]{0.5\linewidth}
    \begin{tikzpicture}[scale=.7, transform shape, 
        level/.style={sibling distance=10mm/#1},
        every edge/.style={draw, thin}
      ]
      \node [circle,draw] (a) {}
      child {node [left=0cm] (b) {$f_1$}}
      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
      child {node [left=.4cm] (e) {$f_n$}}
      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
          child {node [left=.5cm] (b) {$f_1$}} 
          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
          child {node [circle, draw, left=.25cm] (and2) {}
            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
        }
        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
          child {node [circle, draw, right=.25cm] (and3) {}
            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
          child {node [right=.6cm] (e) {$f_n$}}
        }
      };

      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (a1) {} node {} (l);
      \path (d)  edge[very thick]  node[inner sep=0mm,pos=.4] (b1) {} node {} (r);
      \path (a1) edge[bend right] (b1) ;

      \path (and2)  edge node[inner sep=0mm,pos=.4] (t0) {} node {} (l2);
      \path (and2)  edge node[inner sep=0mm,pos=.4] (k0) {} node {} (r2);
      \path (t0) edge[bend right] (k0) ;

      \path (and3)  edge node[inner sep=0mm,pos=.4] (t1) {} node {} (l3);
      \path (and3)  edge node[inner sep=0mm,pos=.4] (k1) {} node {} (r3);
      \path (t1) edge[bend right] (k1) ;
    \end{tikzpicture}
    \caption{  \label{fig:andor}}
    \end{subfigure}
  \caption{(a) The typed combinator $S * I$ ($f(x)=x^2$ over real values ($\mathbb{R}$)) represented as a
    binary tree. (b) The space of all typed combinators represented as an
    infinitely deep AND/OR tree.}

\end{figure}

\subsection{A stochastic grammar over programs}
\label{sec:stochgrammar}

We specify a distribution $\mathcal{D}$ over expressions in
$\mathcal{L}$ as a stochastic
grammar~\cite{feldman1969grammatical}. Many stochastic grammars have
the desirable property that the probability of an expression is the
product of the probabilities of its subcomponents. The distribution we
specify will be a simple version of this, in which the
probability of an expression is the product of the probability of its
leaves.

Let ${C = c_1, \dots, c_N }$ be the primitive combinators in
$\mathcal{L}$. ${ \mathcal{D}=p_1, \dots, p_N }$ will associate with
each primitive combinator $c_i$ a prior probability $p_i$,
$0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $. The probability of a leaf with
primitive combinator $c_i$ will depend on which other primitive
combinators could have been chosen in its place. This set will depend
on the type that was requested by the parent of this leaf when the
leaf's primitive was chosen; we call this type the \emph{requesting}
type of the leaf. For example, suppose that we have the primitive
$(+1)$ of type $Int \rightarrow Int$ and the primitive $NOT$ of type
$Bool \rightarrow Bool$. If the requesting type is $\sigma \rightarrow
Int$, then we can only use $(+1)$, but if the requesting type is
$\sigma \rightarrow \sigma$ then we could use either $(+1)$ or
$NOT$. We define $C_\tau$ for any type $\tau$ to be the set of
primitive combinators in $C$ that are consistent with $\tau$, that is,
whose types \emph{unify} with $\tau$ (we use unification according to
the standard type theory presentation~\cite{Pierce_2002}).

Let $C_e$ be the set of primitive combinators in the leaves of
expressions $e$. We define the probability of an expression ${e \in
  \mathcal{L}}$ to be ${ p(e)~=~\prod_{c \in C_e} p(c | \tau(c)) }$
where $p(c | \tau(c))$ is the probability of using primitive $c$ when
the requesting type is $\tau(c)$. In turn, we define the conditional
probability for each combinator $c_n$ to be $p(c_n| \tau(c_n)) \propto
\frac{p_n}{\sum_{c_j \in C_{\tau(c_n)}} p_j}$.  That is, $p(c_n | \tau(c_n)$ is
proportional to the probability of sampling $c_n$ from the multinomial
distribution defined by the $p_n$'s, conditioned on the requesting
type, with the constant of proportionality chosen to ensure that the
sum over the probabilities of all expressions converges to a finite
value (so that the distribution over expressions is proper). For
binary trees, this is true whenever the constant of proportionality is
less than~$\frac{1}{4}$\footnote{The probability mass $M_d$ of all
  expressions of depth less than or equal to $d$ can be written with
  the recurrence relation $M_d \leq M_{d-1}^2 + M_1$. This has an
  asymptotic fixed point, as $d \rightarrow \infty$ if $x = x^2 + M_1$
  has a solution, which is true if $M_1~\leq~1/4$.}.

Now suppose that we have a set of expressions $\mathcal{E}$, and we
wish to estimate the maximum likelihood values of the $p_n$'s. If the
choice of primitive at each leaf were not conditioned on the
requesting type, we could just count all occurrences of each primitive
in our expression set, and this would be proportional to the maximum
likelihood estimate. This is the ML estimator for a multinomial
distribution. However, the fact that we condition on the requesting
type makes this a considerably more difficult optimization task. One
of the simplest approximations we can make is to calculate the
frequency with which each primitive combinator appears in
$\mathcal{E}$ when the requesting type is such that it could have
been chosen. This is a straightforward calculation, and
we use it for its convenience, finding that our empirical results
justify its use. Future inquiry will be needed to determine to what
extent a more accurate estimator can improve these results.

\subsection{Best-first enumeration of programs}
In order to explore the frontier of most promising programs, we need a
procedure that enumerates the $N$ most probable expressions. There has
been recent interest in this problem, most of it focused on
enumerating the shortest program satisfying some
criterion~\cite{DBLP:conf/sfp/Katayama05,DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
problem as a best-first exploration of an AND/OR tree
(Figure~\ref{fig:andor})~\cite{nilsson1982principles,DBLP:journals/cacm/Hall73}. In
words, the enumeration procedure is best described by the following
recursion: every program is either a primitive combinator OR it is a
left child program applied to a right child program (that is, a left
child AND a right child).

This can framed as the following AND/OR tree $\mathcal{G}$. Suppose we
want a program of type $\tau$. The root of $\mathcal{G}$ is an OR node of
type $\tau$. Its children are elements in $C_\tau$ (defined in the previous
section) and one AND node of type $\tau$ with two children. Each child of
an AND node has the same structure as the root node, with modified
types: the type of its left child is $\sigma \rightarrow \tau$ (where $\sigma$ is
a fresh type variable). Since the type of the right child is
constrained by the subtree rooted at the left child, we always expand
the left child first. Once we have a complete left child program, we
can use its target type as the type of the right child.

Recall that a valid partial solution $\mathcal{H}$ is a subtree of
$\mathcal{G}$ satisfying the following properties: the root of
$\mathcal{H}$ is the root of $\mathcal{G}$ and any OR node in
$\mathcal{H}$ has at most one child. $\mathcal{H}$ is a complete
solution (and thus a complete expression) if it is a partial solution
whose leaves are leaves of $\mathcal{G}$, if each OR node in
$\mathcal{H}$ has exactly one child, and if each AND node in
$\mathcal{H}$ is connected to all of that node's children in
$\mathcal{G}$.

The value of a partial solution $\mathcal{H}$ is calculated by summing
the value of its leaves, where each such value is the log of the
probability assigned to that leaf in
Section~\ref{sec:stochgrammar}. If a leaf of $\mathcal{H}$ is a leaf
of $\mathcal{G}$ then either it is an AND node, to which we assign a
value of 0, or it corresponds to some primitive combinator $c_n$ in
the library and is the child of some typed OR node with type $\tau$. We
assign it a value of $\log{p(c_n | \tau)}$. If a leaf $v$ of
$\mathcal{H}$ is an OR node, then the value of any extension of it to
a complete solution is at most $log(1/4)$. Thus, we have defined a
value for partial solutions that upper bounds the value of any of its
extensions.

Our best-first policy is to expand the next partial solution
with the highest value. If a partial solution has multiple leaves that
are not primitive combinators, we expand the left most one first,
since it is the choice of left combinator that defines the type of its
right combinator in an application pair. That is, the parent node
applies its constraint on valid solutions via the left child, not the
right child. We want to apply constraints as early as possible, so
expanding left child first prunes more partial solutions.

\subsection{Finding the most compressive set of solutions}

Having enumerated the frontier, we want to assign solutions
to hit tasks such that this set is maximally compressible. This will
promote reusable modular subtrees.

>>>>

\documentclass[preview]{standalone}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{subcaption}
\usetikzlibrary{arrows,automata, positioning, patterns}

\begin{document}
\begin{figure}
  \begin{subfigure}{.3\linewidth}
     \begin{tikzpicture} 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 4 ] 4 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.3\linewidth}
     \begin{tikzpicture} 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 5 ] 5 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.3\linewidth}
     \begin{tikzpicture} 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 6 ] 6 ]
     \end{tikzpicture}
  \end{subfigure}

  \begin{subfigure}{.3\linewidth}
     \begin{tikzpicture} 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 4 ] 
     \end{tikzpicture}
  \end{subfigure}
   \begin{subfigure}{.3\linewidth}
     \begin{tikzpicture} 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 5 ] 
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.3\linewidth}
     \begin{tikzpicture} 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 6 ] 
     \end{tikzpicture}
  \end{subfigure}
\end{figure}
